Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-13 algo-best-alt.tex
fin preuve best search

\florent{New Algorithm for best-search}
We shall compute the values of the functions $b_\bot$ and $b_\top$
by search of shortest paths in a $\Semiring$-labeled graph $\G_A$
associated to the \SWVPA $A$. 
This graph is defined by $\G_A = \< V_A, E_A, \eta_A >$, 
with set of vertices $V_A = (Q \times Q) \cup (Q \times P \times Q)$,
set of edges $E_A = V_A \times V_A$
(an edge $\< v_1, v_2>$, with $v_1, v_2 \in V_A$ is denoted by $v_1 \to v_2$),
and with an edge labelling  
$\eta_A : E_A \to \Semiring$ defined by, %as follows,
for $q_0, q_1, q_2, q_3 \in Q$, $p, p' \in P$:
%
\[
\begin{array}{rclcl}
%
\< q_0, q_1> & \to & \< q_0, q_2> & \mapsto &
\bigoplus_{\Deltai} \weiei(q_1, q_2) \oplus \bigoplus_{\Deltar} \weier(q_1, q_2)\\
%
\< q_1, p, q_2> & \to & \<q_0, q_3> & \mapsto &
\bigoplus_{\Deltac} \bigl[ \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr]\\ 
%
%\< q_0, p, q_1> & \to & \< q_0, p, q_2> & : &
%\bigoplus_{\Deltai} \weii(q_1, p, q_2)\\
%
\< q_1, p', q_2> & \to & \< q_0, p, q_3> & \mapsto &
{\displaystyle\bigoplus_{q_0=q_1}}
{\displaystyle\bigoplus_{p=p'}}
 \bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
 & & & & 
\qquad\quad 
\oplus \bigoplus^2_{\Deltac} \bigl[ \weic(q_0, p, p', q_1) \otimes_2 {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
%
\end{array}
\]
%Removing from $\G_A$ every edge $e \in E_A$ such that $\eta(e) = \zero$ 
%is preserving the results below.
%
A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
it is called \emph{safe} when $v_0$ has the form $\< q, q>$ for some $q \in Q$, 
and $n_v$ is called $\last(\pi)$.
%
A path $\pi$ is assigned a weight value $\weight(\pi) \in \Semiring$ which is the product by $\otimes$ 
of the weights of the edges involved in the path.
More precisely, for all $q, q' \in Q$ and $p \in P$, we let 
$\weight(\< q, q' > ) = \one$ if $q = q'$, and $\zero$ otherwise, 
$\weight(\< q, p, q' > ) = \zero$ and 
for all $n \geq 1$, 
$\weight(v_0,\ldots, v_n) = \weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n})$.
%$\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.

%Intuitively, 
\begin{lemma}[Correctness]
For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
%with last element $v$, \\
($i$) if $\last(\pi) = \< q, q'>$, %with $q, q' \in Q$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,\\
%
($ii$) if $\last(\pi) = \< q, p, q'>$, %with $q, q' \in Q$, $p\in P$, 
then there exists $v \in \Delta^*$ such that \\
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length $|\pi|$ of the path $\pi$.

@@

\begin{lemma}[Completeness] \label{lem:algo-completeness}
For all $q, q' \in Q$, $p\in P$, 
($i$) there exists a safe path~$\pi$ of~$\G_A$ such that
%such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\last(\pi) = \< q, q'>$,
and $\weight(\pi) = b_\bot(q, q')$, 
($ii$) there exists a safe path $\pi'$ of $\G_A$ 
such that $\last(\pi') = \< q, p, q'>$,
and $\weight(\pi') = \bigoplus_\Deltac b_\top(q, p, q')$.
\end{lemma}
%
\begin{proof}
By associativity, commutativity and distributivity for $\Semiring$,
\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
%
\begin{equation}\label{eq:bbot-unfold}
b_\bot(q, q') = 
\bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
where $n$ is the length of $t$, $k \leq n$,
$q_0 = q$, $q_n = q'$, 
for all $1 \leq i < n$, 
$w_i$ is one of the functions of $\bar\wei$,
$\tau_i$ is a transition of $A$ and
$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
%
Since $\Semiring$ is total, there exists finite sequences as above such that
\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
%
There might exists, for $q$ and $q'$,
several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
let us choose arbitrarily one of minimal length~$n$. 
This integer $n$ will be denoted $n_{q, q'}$ in the following.
%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,

@@

\noindent
For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
be the sequences associated to $q$ and $q'$ as above.
Let us do a case analysis of $w_n$.

\end{proof}

>>>>

\florent{New Algorithm for best-search}
We shall compute the values of the functions $b_\bot$ and $b_\top$
by search of shortest paths in a $\Semiring$-labeled graph $\G_A$
associated to the \SWVPA $A$. 
This graph is defined by $\G_A = \< V_A, E_A, \eta_A >$, 
with set of vertices $V_A = (Q \times Q) \cup (Q \times P \times Q)$,
set of edges $E_A = V_A \times V_A$
(an edge $\< v_1, v_2>$, with $v_1, v_2 \in V_A$ is denoted by $v_1 \to v_2$),
and with an edge labelling  
$\eta_A : E_A \to \Semiring$ defined by, %as follows,
for $q_0, q_1, q_2, q_3 \in Q$, $p, p' \in P$:
%
\[
\begin{array}{rclcl}
%
\< q_0, q_1> & \to & \< q_0, q_2> & \mapsto &
\bigoplus_{\Deltai} \weiei(q_1, q_2) \oplus \bigoplus_{\Deltar} \weier(q_1, q_2)\\
%
\< q_1, p, q_2> & \to & \<q_0, q_3> & \mapsto &
\bigoplus_{\Deltac} \bigl[ \weiec(q_0, p, q_1) \otimes \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr]\\ 
%
%\< q_0, p, q_1> & \to & \< q_0, p, q_2> & : &
%\bigoplus_{\Deltai} \weii(q_1, p, q_2)\\
%
\< q_1, p', q_2> & \to & \< q_0, p, q_3> & \mapsto &
{\displaystyle\bigoplus_{q_0=q_1}}
{\displaystyle\bigoplus_{p=p'}}
 \bigoplus_{\Deltai} \weii(q_2, p, q_3)\\
 & & & & 
\qquad\quad 
\oplus \bigoplus^2_{\Deltac} \bigl[ \weic(q_0, p, p', q_1) \otimes_2 {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\\
%
\end{array}
\]
%Removing from $\G_A$ every edge $e \in E_A$ such that $\eta(e) = \zero$ 
%is preserving the results below.
%
A \emph{path} of $\G_A$ is a sequence $\pi = v_0,\ldots, v_n \in V^*_A$,
where $v_0$ and $v_n$ are respectively called $\fst(\pi)$ and $\last(\pi)$.
The path $\pi$ is called \emph{safe} when $\fst(\pi)$ has the form $\< q, q>$ for some $q \in Q$, 
%
A path $\pi$ is assigned a weight value $\weight(\pi) \in \Semiring$ which is the product by $\otimes$ 
of the weights of the edges involved in the path.
More precisely, for all $q, q' \in Q$ and $p \in P$, we let 
$\weight(\< q, q' > ) = \one$ if $q = q'$, and $\zero$ otherwise, 
$\weight(\< q, p, q' > ) = \zero$ and 
for all $n \geq 1$, 
$\weight(v_0,\ldots, v_n) = \weight(v_0,\ldots, v_{n-1}) \otimes \eta_A(v_{n-1} \to v_{n})$.
%$\weight(\pi) = \bigotimes_{i=1}^{n} \eta(v_{i-1} \to v_i)$.
%
%The following results establish that $b_\bot(q, q')$ by be computed from a shortest path in $\G_A$.
For $v, v' \in V_A$, let $\mathit{short}(v, v')$ be the weight of the shortest path from $v$ to $v'$, 
more precisely, 
\(\mathit{short}(v, v') = \displaystyle\bigoplus_{\begin{array}{c}
                                                  \scriptstyle\pi\,\mathrm{s.t.}\,
										          \scriptstyle\fst(\pi) = v,
	                                              \scriptstyle\last(\pi) = v'
											      \end{array}} \weight(\pi)$.

\begin{proposition}\label{algo-shortest}
%Let  and let $\pi$ be the shortest safe path with $\last(\pi) = \< q, q'>$, i.e. 
For all $q, q' \in Q$, 
$b_\bot(q, q') = \displaystyle\bigoplus_{q_0 \in Q} \mathit{short}\bigl(\< q_0, q_0>, \< q, q'>\bigr)$.
\end{proposition}
This proposition is a consequence of Lemmata~\ref{algo-correct} and~\ref{algo-complete} below.

%Intuitively, 
\begin{lemma}[Correctness]\label{algo-correct}
For all safe path~$\pi$ of~$\G_A$,  %with $\src(\pi) = \<q_0, q_0>$,
%with last element $v$, \\
($i$) if $\last(\pi) = \< q, q'>$, %with $q, q' \in Q$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,\\
%
($ii$) if $\last(\pi) = \< q, p, q'>$, %with $q, q' \in Q$, $p\in P$, 
then there exists $v \in \Delta^*$ such that \\
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length $|\pi|$ of the path $\pi$.

@@

\begin{lemma}[Completeness] \label{algo-complete}
For all $q, q' \in Q$, $p\in P$, 
($i$) there exists a safe path~$\pi$ of~$\G_A$ such that
%such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\last(\pi) = \< q, q'>$,
and $\weight(\pi) = b_\bot(q, q')$, 
($ii$) there exists a safe path $\pi'$ of $\G_A$ 
such that $\last(\pi') = \< q, p, q'>$,
and $\weight(\pi') = \bigoplus_\Deltac b_\top(q, p, q')$.
\end{lemma}
%
\begin{proof}
By associativity, commutativity and distributivity for $\Semiring$,
\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
%
\begin{equation}\label{eq:bbot-unfold}
b_\bot(q, q') = 
\bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
where $n$ is the length of $t$, $k \leq n$,
$q_0 = q$, $q_n = q'$, 
for all $1 \leq i < n$, 
$w_i$ is one of the functions of $\bar\wei$,
$\tau_i$ is a transition of $A$ and
$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
%
Since $\Semiring$ is total, there exists finite sequences as above such that
\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
%
There might exists, for $q$ and $q'$,
several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
let us choose arbitrarily one of minimal length~$n$. 
This integer $n$ will be denoted $n_{q, q'}$ in the following.
%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,

@@

\noindent
For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
be the sequences associated to $q$ and $q'$ as above.
Let us do a case analysis of $w_n$.
... \qed
\end{proof}

\medskip\noindent
Following~\eqref{eq:min} and Proposition~\ref{algo-shortest}, 
in order to compute the minimum of $t \mapsto A(t)$ \wrt $\leq_\oplus$, 
it is sufficient to compute one shortest path in $\G_{A}$ of source
$\<q_0, q_0>$ and target $\<q, q'>$ for all $q_0, q, q' \in Q$.
This can be done with~~\cite{Mohri02semiring}~Theorem~?? \florent{completer}
with a complexity ***. Altogether, the complexity for computing the minimum of $A$
is in $O(...)$.
Moreover, a witness $t \in \Delta^*$ for this minimum can be associated to the appropriate
shortest path, like in the proof of Lemma~\ref{algo-correct}

### 2021-10-13 introduction.tex
fin preuve best search

\begin{figure}
\centering
\begin{tikzpicture}
\node (SWADS) at (0,5.0) {%
  \(
  \begin{array}{c}
  \mathsf{SWADS}\,\mbox{\cite{Herrmann16dlt,Herrmann20phd}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}
  \end{array}
  \)
};
%
\node (SWA) at (0,3.4) {%
  \(
  \begin{array}{c}
  \mathsf{SWA}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
  \end{array}
  \)
};
%
\node (WA) at (-1.8,1.7) {%
  \(
  \begin{array}{c}
  \mathsf{WA}\,\mbox{\cite{Droste09handbook}}: \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
  q \xrightarrow{a, w} q', a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
  \end{array}
  \)
};
%
\node (SA) at (1.8,1.7) {%8
  \(
  \begin{array}{c}
  \mathsf{SA}\,\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
  \end{array}
  \)
};
%
\node (NFA) at (0,0) {%
  \(
  \begin{array}{c}
  \mathsf{FA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
  q \xrightarrow{a} q', a \in \Sigma_\mathsf{fin}
  \end{array}
  \)
};
%
\node (VPA) at (5.6,0) {%
  \(
  \mathsf{VPA}\,\cite{AlurMadhusudan09nested}
  \)
};
%
\node (WVPA) at (4.6,1.7) {%
  \(
  \mathsf{WVPA}\,\cite{Caralp12VPAmult}
  \)
};
%
\node (SVPA) at (6.6,1.7) {%
  \(
  \mathsf{SVPA}\,\cite{dAntonyAlur14SVPDA}
  \)
};
%
\node (SWVPA) at (5.6,3.4) {%
  \(
  \mathsf{SWVPA}
  \)
};
%
\draw[->] (NFA)--(WA);
\draw[->] (NFA)--(SA);
\draw[->] (WA)--(SWA);
\draw[->] (SA)--(SWA);
\draw[->] (SWA)--(SWADS);
\draw[->] (VPA)--(WVPA);
\draw[->] (VPA)--(SVPA);
\draw[->] (WVPA)--(SWVPA);
\draw[->] (SVPA)--(SWVPA);
\draw[->] (SWVPA)--(SWADS);
%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array}
\end{tikzpicture}
\caption{Classes of Symbolic/Weighted Automata.
Here, $\Sigma_\mathsf{fin}$ and $\Sigma_\mathsf{inf}$ denote finite/countable alphabets,
$\mathbb{B}$  the Boolean algebra,
$\mathbb{S}$ a commutative semiring.
$q \xrightarrow{\dots} q'$ is a transition between states $q$ and $q'$.}
\label{fig:hierarchy}
\end{figure}
%
%This approach is close to the case of
%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
%except that the domain for weight values is not restricted to be Boolean,
%like for the guards in the rules of SA,
%but can be an arbitrary commutative semiring (assuming some restrictions).

@@

The output of parsing is a sequence of
timed symbols $\nu\tstamp{\tau'}$ in an alphabet $\Delta$,
%events (or \emph{notes})
%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where $\nu$  represents
an \emph{event} (or \emph{note}),
specified by its \emph{pitch name}
(\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.),
an event \emph{continuation} (symbol `$-$`, see Example~\ref{ex:SWT}),
or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
The temporal information $\tau'$ 
is either a time interval, for the opening parentheses
(representing the duration between the parenthesis 
 and the matching closing one), 
or a timestamp, for the other symbols.
The time points in $\tau'$ belong to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) split in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), \etc.
%
For instance, the output score
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
  %\includegraphics[scale=0.20]{pictures/score5.png},
corresponds to a hierarchical structure
that can be linearized as the sequence
$O =$
$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
$\ccall{2}\tstamp{[0,1]}$,
$\mbox{\footnotesize A4}\tstamp{0}$,
$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
$-\tstamp{\frac{1}{2}}$,
$\ccall{2}\tstamp{[\frac{3}{4},1]}$,
$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
$\mbox{\footnotesize C$\sharp$5}\tstamp{\frac{7}{8}}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
$\ccall{3}\tstamp{[1,2]}$,
$\mbox{\footnotesize D5}\tstamp{1}$,
$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
$\creturn{3}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$.
%\end{enumerate}
The opening markups $\ccall{\mathsf{m}}$ %and $\creturn{\mathsf{m}}$
delimit \emph{measures},
which are time intervals of duration 1 in this example,
while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,
for some natural number~$d$,
represent a division of the time interval attached to $\ccall{d}$,
of duration $\ell$,
into $d$ sub-intervals of equal duration $\frac{\ell}{d}$.
%
We will show that $O$ is a solution for the
parsing of $I$. Note that several other parsings are possible
like \eg \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}

>>>>

\begin{figure}
\centering
\begin{tikzpicture}
\node (SWADS) at (3.0,5.0) {%
  \(
  \begin{array}{c}
  \mathsf{SWADS}\,\mbox{\cite{Herrmann16dlt,Herrmann20phd}}
  \end{array}
  \)
};
%
\node (SWA) at (0,3.4) {%
  \(
  \begin{array}{c}
  \mathsf{SWA}: \Sigma_{\mathsf{inf}}^* \to \mathbb{S}\\
  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{S}
  \end{array}
  \)
};
%
\node (WA) at (-1.8,1.7) {%
  \(
  \begin{array}{c}
  \mathsf{WA}\,\mbox{\cite{Droste09handbook}}: \Sigma_{\mathsf{fin}}^* \to \mathbb{S}\\
  q \xrightarrow{a, w} q', a \in \Sigma_\mathsf{fin}, w \in \mathbb{S}
  \end{array}
  \)
};
%
\node (SA) at (1.8,1.7) {%8
  \(
  \begin{array}{c}
  \mathsf{SA}\,\mbox{\cite{dAntoni21CACM}}: \Sigma_{\mathsf{inf}}^* \to \mathbb{B}\\
  q \xrightarrow{\phi} q', \phi : \Sigma_\mathsf{inf} \to \mathbb{B}
  \end{array}
  \)
};
%
\node (NFA) at (0,0) {%
  \(
  \begin{array}{c}
  \mathsf{FA} : \Sigma_{\mathsf{fin}}^* \to \mathbb{B}\\
  q \xrightarrow{a} q', a \in \Sigma_\mathsf{fin}
  \end{array}
  \)
};
%
\node (VPA) at (5.6,0) {%
  \(
  \mathsf{VPA}\,\mbox{\cite{AlurMadhusudan09nested}}
  \)
};
%
\node (WVPA) at (4.6,1.7) {%
  \(
  \mathsf{WVPA}\,\mbox{\cite{Caralp12VPAmult}}
  \)
};
%
\node (SVPA) at (6.6,1.7) {%
  \(
  \mathsf{SVPA}\,\mbox{\cite{dAntonyAlur14SVPDA}}
  \)
};
%
\node (SWVPA) at (5.6,3.4) {%
  \(
  \mathsf{SWVPA}
  \)
};
%
\draw[->] (NFA)--(WA);
\draw[->] (NFA)--(SA);
\draw[->] (WA)--(SWA);
\draw[->] (SA)--(SWA);
\draw[->] (SWA)--(SWADS);
\draw[->] (VPA)--(WVPA);
\draw[->] (VPA)--(SVPA);
\draw[->] (WVPA)--(SWVPA);
\draw[->] (SVPA)--(SWVPA);
\draw[->] (SWVPA)--(SWADS);
%\begin{array}{c} \mathsf{NFA} : \Sigma^* \to \mathbb{B} \end{array}
\end{tikzpicture}
\caption{Classes of Symbolic/Weighted Automata.
Here, $\Sigma_\mathsf{fin}$ and $\Sigma_\mathsf{inf}$ denote finite/countable alphabets,
$\mathbb{B}$  the Boolean algebra,
$\mathbb{S}$ a commutative semiring.
$q \xrightarrow{\dots} q'$ is a transition between states $q$ and $q'$.}
\label{fig:hierarchy}
\end{figure}
%
%This approach is close to the case of
%Symbolic Automata (SA)~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
%except that the domain for weight values is not restricted to be Boolean,
%like for the guards in the rules of SA,
%but can be an arbitrary commutative semiring (assuming some restrictions).

@@

The output of parsing is a sequence of
timed symbols $\nu\tstamp{\tau'}$ in an alphabet $\Delta$,
%events (or \emph{notes})
%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where $\nu$  represents
an \emph{event} (or \emph{note}),
specified by its \emph{pitch name}
(\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.),
an event \emph{continuation} (symbol `$-$`, see Example~\ref{ex:SWT}),
or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
The temporal information $\tau'$ 
is either a time interval, for the opening parentheses
(representing the duration between the parenthesis 
 and the matching closing one), 
or a timestamp, for the other symbols.
The time points in $\tau'$ belong to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) split in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), \etc.
%
For instance, the output score
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
  %\includegraphics[scale=0.20]{pictures/score5.png},
corresponds to a hierarchical structure
that can be linearized as the sequence
$O =$
$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
$\ccall{2}\tstamp{[0,1]}$,
$\mbox{\footnotesize A4}\tstamp{0}$,
$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
$-\tstamp{\frac{1}{2}}$,
$\ccall{2}\tstamp{[\frac{3}{4},1]}$,
$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
$\mbox{\footnotesize C$\sharp$5}\tstamp{\frac{7}{8}}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\creturn{\mathsf{m}}\tstamp{1}$,
$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
$\ccall{3}\tstamp{[1,2]}$,
$\mbox{\footnotesize D5}\tstamp{1}$,
$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
$\creturn{3}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$.
%\end{enumerate}
The opening markups $\ccall{\mathsf{m}}$ %and $\creturn{\mathsf{m}}$
delimit \emph{measures},
which are time intervals of duration~1 in this example,
while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,
for some natural number~$d$,
represent a division of the time interval attached to $\ccall{d}$,
of duration $\ell$,
into $d$ sub-intervals of equal duration $\frac{\ell}{d}$.
%
We will show that $O$ is a solution for the
parsing of $I$. Note that several other parsings are possible
like \eg \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}
